Turbulence in compact to giant H~II regions

### 2023-04-26 strucfunc_paper.tex
Incomplete edits on NGC 595 and 606

NGC 346 is the most active star-formation region in the SMC (Small Magellanic Cloud) located at a distance of 62 kpc \citetext{\SI{1}{\arcsecond} = \SI{0.30}{pc} ; \citealp{2001ApJ...562..303D}}. 
It has more than 30 O stars that ionize N66, the largest \hii{} region \citep{2011ApJ...740...10D}.
\NEW{The cluster}
has an ionizing output \(Q_0\) of \SI{e50.71}{photons.s^{-1}} (40 times higher than the Orion Nebula) and \NEW{the \hii{} region has} an \ha{} luminosity of \SI{e38.77}{erg.s^{-1}} \citep{2010A&A...517A..39H,1984ApJ...287..116K}.
\startNEW
A supernova remnant is present in the region \citetext{J59.4\(-7210\), \citealp{Ye:1991d}}
but it does not overlap with the MUSE field that we observe here \citep{Maggi:2019q}.
High-resolution \ha{} spectroscopy of the \hii{} region
gives a non-thermal velocity width of \(\sigma\los = \SI{10.1}{km.s^{-1}}\)
\citep{2003ApJ...586.1179D}.
\stopNEW
\citet{2008ApJ...688.1050G} propose an expanding \hii{} region or bubble blown by the winds of the massive progenitor as a mechanism that shapes the recent star formation in this region, in addition to the photoionization from the OB stars. 
This mechanism is similar to shell-like \hii{} regions, with a central cluster in a cavity, and with ongoing star formation triggered around their periphery.

@@

Hubble V is located near the ``Hodge OB 8'' OB association with a number of \num{\sim 36} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O3 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.23}{photons.s^{-1}} (13 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.30}{erg.s^{-1}}\citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} \(\sigma\los = \SI{10.3 \pm 0.03}{km.s^{-1}}\) \citep{1986A&A...160..374H}.

Hubble X is located near the Hodge ``OB 13'' OB association with a number of \num{\sim 29} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O4 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.14}{photons.s^{-1}} (10 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.21}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \(\sigma\los = \SI{10.5 \pm 0.02}{km.s^{-1}}\)  \citep{1986A&A...160..374H}.
\citet{1993PASJ...45..693T} studies the kinematics of both regions.

@@

In the Figure~\ref{fig:strucfunc-fit-ExtraGal}a we show the structure function of Hubble V.
\startNEW
The model-derived parameters are \(\sigma\pos^2 = \SI{10 \pm 3}{km^{2}.s^{-2}}\),
\(r_0 = \SI{3.6 \pm 1.0}{pc} \)), and \(m = 0.8 \pm 0.3\).
The velocity variance is slightly higher than the empirically derived value of \(\sigma\pos^2 = \SI{8}{km^{2}.s^{-2}}\) because the model-derived seeing is only 6 times smaller than the correlation length.

@@

NGC 595 is the second most luminous giant \hii{} region in the M33 galaxy,
\startNEW
excited by a stellar population comprising
250 OB stars, 13 hot supergiants, and 11 Wolf-Rayet stars \citep{1993AJ....105.1400D, 1996AJ....111.1128M}.
\stopNEW
The ionizing output \(Q_0\) is \SI{e50.88}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and
the \hii{} region has a luminosity in \ha{} of \SI{e38.95}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\sigma\los = \SI{16.8 \pm 0.1}{km.s^{-1}}\) \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find an \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the
\startNEW
most luminous \hii{} region in M33 and
\stopNEW
the second most luminous in the Local Group after 30~Dor.
\startNEW
The dominant stellar cluster is located in the center of the region and contains 200 OB stars and 4 Wolf-Rayet stars \citep{1996ApJ...456..174H, 2011MNRAS.411..235E}.
The optically visible stars come from a population
with an age of \SI{4}{Myr},
although infrared observations detect the presence of
ongoing massive star formation in the region
\citep{2012AJ....143...43F, 2012ApJ...761....3M}.
\stopNEW
The ionizing output is \(Q_0 = \SI{e51.35}{photons.s^{-1}}\)
(175 times higher than the Orion Nebula) and
the region has an \ha{} luminosity of \SI{e39.42}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\sigma\los = \SI{17.8 \pm 0.3}{km.s^{-1}}\) \citep{1986A&A...160..374H}.
\startNEW
The morphology of the region is strikingly similar to that of 30~Dor
(see Figure~\ref{fig:hii-regions})
\stopNEW
and is visually dominated by a predominant big loop, many expanding shells and different size filaments.

@@

In the Figure~\ref{fig:strucfunc-fit-ExtraGal}c we show the structure function of NGC 595.
The value of the observed \(\sigma\pos^2 = \SI{44}{km^{2}.s^{-2}}\) is obtained at \(r_0 = \SI{10}{pc} \) (while the derived values from our model are \(\sigma\pos^2 = \SI{53}{km^{2}.s^{-2}}\) and \(r_0 = \SI{11}{pc} \)).
The observed value of \(2\sigma\pos^2\) corresponds to \(r = \SI{21}{pc} \).
The maximum \(B(r)\) value is \SI{104}{km^{2}.s^{-2}} at a separation of \SI{30}{pc}. 
%For NGC 595 we see that the structure function plateau at this maximum \(B(r)\) value, which implies that the observed \(\sigma\pos^2\) is affected by the observational constraints.
Until \(r_0 \sim \SI{55}{pc} \) the structure functions remains closely to the 'true' \(2\sigma\pos^2\), for the largest scale there is a decrease between  \(\SI{55}{pc} < r < \SI{110}{pc} \) where the structure function plateaus at the 'observed' \(2\sigma\pos^2\). 
Above \SI{110}{pc} we see a very disperse downturn  until \SI{54}{km^{2}.s^{-2}} at a separation of \SI{211}{pc}.
The seeing (RMS) is \SI{2.1}{pc} (\(s_0\)=\SI{0.2}{\arcsecond}) and there are no points below this value.
Considering the first point to be valid, which corresponds to \SI{2.5}{km^{2}.s^{-2}} at \(r = \SI{1.05}{pc} \) the slope for \(\sigma\pos^2\) and \(2\sigma\pos^2\) we obtain \num{1.27} and \num{1.18}, respectively, while our derived value is \num{1.38}.

@@

\subsubsection{Comparison with previous studies}
\label{sec:comparison-m33}
For our NGC 595 \halpha\ emission line results there is no correspondence in the structure function from the previous investigations of \citet{lagrois2009multi} and \citet{lagrois2011}.
Our observations cover the brightest part of the region and have smaller resolution than theirs.
As a consequence \citet{lagrois2009multi} and \citet{lagrois2011} structure functions results does not cover scales $<$10 pc, where in our results the correlation is taking place.
One big difference is that \citet{lagrois2011} apply a  filter to the velocity field to get rid of non-turbulent movements.
As a comparison, for the filtered turbulent parameters they chose to represent the turbulent field, the derive a total  decorrelation length (\(C(\tau_0) = 0\)) of \SI{43}{pc} and the power-law index they provide is \num{1.55 \pm 0.01}.
Their velocity centroid \(\sigma\) is \SI{5.92 \pm 0.15}{km.s^{-1}} while we have \SI{7.3\pm 2}{km. s^{-1}} which fall within the uncertainties.
Due to the different observations and methodologies a direct comparison between structure functions is not possible.

@@

\section{Additional covariance corner plots for model fits}
\label{sec:addit-covar-corn}
Supplementary material
\href{https://github.com/will-henney/javier-strucfunc-paper/blob/main/supplementary_material_strucfunc_paper.pdf}
{available online}.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% END %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

>>>>

NGC 346 is the most active star-formation region in the SMC (Small Magellanic Cloud) located at a distance of 62 kpc \citetext{\SI{1}{\arcsecond} = \SI{0.30}{pc} ; \citealp{2001ApJ...562..303D}}. 
It has more than 30 O stars that ionize N66, the largest \hii{} region \citep{2011ApJ...740...10D}.
\NEW{The cluster}
has an ionizing output \(Q_0\) of \SI{e50.71}{photons.s^{-1}} (40 times higher than the Orion Nebula) and \NEW{the \hii{} region has} an \ha{} luminosity of \SI{e38.77}{erg.s^{-1}} \citep{2010A&A...517A..39H,1984ApJ...287..116K}.
\startNEW
A supernova remnant is present in the region \citetext{J59.4\(-7210\), \citealp{Ye:1991d}}
but it does not overlap with the MUSE field that we observe here \citep{Maggi:2019q}.
High-resolution \ha{} spectroscopy of the \hii{} region
gives a non-thermal velocity width of \(\siglosm = \SI{10.1}{km.s^{-1}}\)
\citep{2003ApJ...586.1179D}.
\stopNEW
\citet{2008ApJ...688.1050G} propose an expanding \hii{} region or bubble blown by the winds of the massive progenitor as a mechanism that shapes the recent star formation in this region, in addition to the photoionization from the OB stars. 
This mechanism is similar to shell-like \hii{} regions, with a central cluster in a cavity, and with ongoing star formation triggered around their periphery.

@@

Hubble V is located near the ``Hodge OB 8'' OB association with a number of \num{\sim 36} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O3 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.23}{photons.s^{-1}} (13 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.30}{erg.s^{-1}}\citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} \(\siglosm = \SI{10.3 \pm 0.03}{km.s^{-1}}\) \citep{1986A&A...160..374H}.

Hubble X is located near the Hodge ``OB 13'' OB association with a number of \num{\sim 29} young stellar objects \citep{2021MNRAS.507.5106K} and its core is powered by an O4 V star \citep{1999PASP..111.1382O}.
It has an ionizing output \(Q_0\) of \SI{e50.14}{photons.s^{-1}} (10 times higher compared to the Orion Nebula) and has a \ha{} luminosity of \SI{e38.21}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is \(\siglosm = \SI{10.5 \pm 0.02}{km.s^{-1}}\)  \citep{1986A&A...160..374H}.
\citet{1993PASJ...45..693T} studies the kinematics of both regions.

@@

In Figure~\ref{fig:strucfunc-fit-ExtraGal}a we show the structure function of Hubble V.
\startNEW
The model-derived parameters are \(\sigma\pos^2 = \SI{10 \pm 3}{km^{2}.s^{-2}}\),
\(r_0 = \SI{3.6 \pm 1.0}{pc} \)), and \(m = 0.8 \pm 0.3\).
The velocity variance is slightly higher than the empirically derived value of \(\sigma\pos^2 = \SI{8}{km^{2}.s^{-2}}\) because the model-derived seeing is only 6 times smaller than the correlation length.

@@

NGC 595 is the second most luminous giant \hii{} region in the M33 galaxy,
\startNEW
excited by a stellar population comprising
250 OB stars, 13 hot supergiants, and 11 Wolf-Rayet stars \citep{1993AJ....105.1400D, 1996AJ....111.1128M}.
\stopNEW
The ionizing output \(Q_0\) is \SI{e50.88}{photons.s^{-1}} (60 times higher compared to the Orion Nebula) and
the \hii{} region has a luminosity in \ha{} of \SI{e38.95}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{16.8 \pm 0.1}{km.s^{-1}}\) \citep{lagrois2009multi}.
\citet{lagrois2011} present a kinematic study of the regions using optical emission lines and and radio 21 cm line observations and find an \ha{} shell morphology that is related to the stellar winds of the several massive stars located in its interior.
They proposed champagne flows at the periphery of the molecular cloud leading to accelerated ionized material.

@@

NGC 604 is the
\startNEW
most luminous \hii{} region in M33 and
\stopNEW
the second most luminous in the Local Group after 30~Dor.
\startNEW
The dominant stellar cluster is located in the center of the region and contains 200 OB stars and 4 Wolf-Rayet stars \citep{1996ApJ...456..174H, 2011MNRAS.411..235E}.
The optically visible stars come from a population
with an age of \SI{4}{Myr},
although infrared observations detect the presence of
ongoing massive star formation in the region
\citep{2012AJ....143...43F, 2012ApJ...761....3M}.
\stopNEW
The ionizing output is \(Q_0 = \SI{e51.35}{photons.s^{-1}}\)
(175 times higher than the Orion Nebula) and
the region has an \ha{} luminosity of \SI{e39.42}{erg.s^{-1}} \citep{2002MNRAS.329..481B}.
The mean non-thermal linewidth in \ha{} is
\(\siglosm = \SI{17.8 \pm 0.3}{km.s^{-1}}\) \citep{1986A&A...160..374H}.
\startNEW
The morphology of the region is strikingly similar to that of 30~Dor
(see Figure~\ref{fig:hii-regions})
\stopNEW
and is visually dominated by a predominant big loop, many expanding shells and different size filaments.

@@

In Figure~\ref{fig:strucfunc-fit-ExtraGal}c we show the structure function of NGC 595.
The model-derived parameters are \(\sigma\pos^2 = \SI{53 \pm 5}{km^{2}.s^{-2}}\),
\(r_0 = \SI{11 \pm 1.0}{pc} \)), and \(m = 1.4 \pm 0.2\).
The value of the observed \(\sigma\pos^2 = \SI{44}{km^{2}.s^{-2}}\) is obtained at \(r_0 = \SI{10}{pc} \) (while the derived values from our model are \(\sigma\pos^2 = \SI{53}{km^{2}.s^{-2}}\) and \(r_0 = \SI{11}{pc} \)).
The observed value of \(2\sigma\pos^2\) corresponds to \(r = \SI{21}{pc} \).
The maximum \(B(r)\) value is \SI{104}{km^{2}.s^{-2}} at a separation of \SI{30}{pc}. 
%For NGC 595 we see that the structure function plateau at this maximum \(B(r)\) value, which implies that the observed \(\sigma\pos^2\) is affected by the observational constraints.
Until \(r_0 \sim \SI{55}{pc} \) the structure functions remains closely to the 'true' \(2\sigma\pos^2\), for the largest scale there is a decrease between  \(\SI{55}{pc} < r < \SI{110}{pc} \) where the structure function plateaus at the 'observed' \(2\sigma\pos^2\). 
Above \SI{110}{pc} we see a very disperse downturn  until \SI{54}{km^{2}.s^{-2}} at a separation of \SI{211}{pc}.
The seeing (RMS) is \SI{2.1}{pc} (\(s_0\)=\SI{0.2}{\arcsecond}) and there are no points below this value.
Considering the first point to be valid, which corresponds to \SI{2.5}{km^{2}.s^{-2}} at \(r = \SI{1.05}{pc} \) the slope for \(\sigma\pos^2\) and \(2\sigma\pos^2\) we obtain \num{1.27} and \num{1.18}, respectively, while our derived value is \num{1.38}.

@@

\startNEW
\subsubsection{Comparison with previous studies}
\label{sec:comparison-m33}

\citet{lagrois2011} calculated the structure function for the \halpha\ emission line for NGC 595
over a region of diameter \SI{300}{kpc} with a resolution of \SI{6.6}{pc.pixel^{-1}} \citep{lagrois2009multi}, presenting results for separations \(r \ge \SI{16.5}{pc}\).
They found a total velocity variance of \(\sigma\pos^2 = \SI{35 \pm 2}{km^{2}.s^{-2}}\), which is smaller than our derived value of \(\sigma\pos^2 = \SI{53}{km^{2}.s^{-2}}\).

For our NGC 595 \halpha\ emission line results there is no correspondence in the structure function from the previous investigations of \citet{lagrois2009multi} and \citet{lagrois2011}.
Our observations cover the brightest part of the region and have smaller resolution than theirs.
As a consequence \citet{lagrois2009multi} and \citet{lagrois2011} structure functions results does not cover scales $<$10 pc, where in our results the correlation is taking place.
One big difference is that \citet{lagrois2011} apply a  filter to the velocity field to get rid of non-turbulent movements.
As a comparison, for the filtered turbulent parameters they chose to represent the turbulent field, the derive a total  decorrelation length (\(C(\tau_0) = 0\)) of \SI{43}{pc} and the power-law index they provide is \num{1.55 \pm 0.01}.
Their velocity centroid \(\sigma\) is \SI{5.92 \pm 0.15}{km.s^{-1}} while we have \SI{7.3\pm 2}{km. s^{-1}} which fall within the uncertainties.
Due to the different observations and methodologies a direct comparison between structure functions is not possible.

@@

\stopNEW

\section{Additional covariance corner plots for model fits}
\label{sec:addit-covar-corn}
Supplementary material
\href{https://github.com/will-henney/javier-strucfunc-paper/blob/main/supplementary_material_strucfunc_paper.pdf}
{available online}.

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% END %%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}
